---
title: "01_exposure_assmt"
author: "Jagadeesh Puvvula"
date: "2025-01-01"
output: pdf_document
---

#pull data from PEACH/CRIB heat-metabolome project
```{r}
df<- read_csv(paste0(analy_dat, "df_for_exposures.csv")) |>
  dplyr::select(-c(3:7, 9, 11:16, 18:20, 22, 33:37, 40)) |>
  mutate(lmp_date = as.Date(mdy(lmp_date), format="%Y-%m-%d"),
         delivery_date = as.Date(mdy(delivery_date), format="%Y-%m-%d")) |>
  drop_na(9, 18) |>
  rename(id=pt_id, lat= latitude, long = longitude) 
```

#want to assign exposures during third trimester (190 to 266 days into pregnancy)
#has 193 participants with complete data, dropped to 177 due to unavialbility of exposure data
```{r}
df_exp<- df |>
  mutate(start_date = lmp_date + days(190),
         end_date = lmp_date + days(266)) |>
  select(id, start_date, end_date, lat, long) |>
  filter(end_date < as.Date("2020-01-01"))
```

#assign particulate matter and black carbon exposures
#Need to update the filename pattern (USHAP_BC or USHAP_PM2.5) in the function
```{r}
bc<- getRasterValueInRange(dataframe = df_exp,
                      input_folder_name = "~/Documents/air_polln_rna_seq/BC_1km_2000_2020",
                      batch_size = 10)

pm<- getRasterValueInRange(dataframe = df_exp,
                      input_folder_name = "~/Documents/air_polln_rna_seq/PM_1km_2000_2020",
                      batch_size = 10)
```

#clean exposure data
```{r}
pm_clean <- pm |>
  mutate(
    raster_value = map(raster_value, ~ as.numeric(.x)),
    pm = map(raster_value, ~ .x[1:length(.x)]), # Extract all values
    pm_expanded = map(pm, ~ set_names(.x, paste0("pm_", seq_along(.x)))) # Add unique names
  ) |>
  select(-raster_value) |> 
  unnest_wider(pm_expanded) |> # Unnest pm_expanded into multiple columns
  mutate(across(starts_with("pm_"), as.numeric)) |> # Ensure new columns are numeric
  select(-c(2:6))

bc_clean <- bc |>
  mutate(
    raster_value = map(raster_value, ~ as.numeric(.x)),
    bc = map(raster_value, ~ .x[1:length(.x)]), # Extract all values
    bc_expanded = map(bc, ~ set_names(.x, paste0("bc_", seq_along(.x)))) # Add unique names
  ) |>
  select(-raster_value) |> 
  unnest_wider(bc_expanded) |> # Unnest pm_expanded into multiple columns
  mutate(across(starts_with("bc_"), as.numeric)) |> # Ensure new columns are numeric 
  select(-c(2:6))

ap_exposures<- left_join(pm_clean, bc_clean, by="id")

save(bc, bc_clean, pm, pm_clean, ap_exposures, file = "~/Documents/air_polln_metabolomics/analytic_dat/exposure_assmnt/exposure_data.rda")
```

#join exposure to analytic data
```{r}
#add linkage variable 
df_join<- df |>select(c(23,9)) |>clean_names() |>rename(id= pt_id, exp_id = client_sample_number)|>
  mutate(exp_id = as.character(exp_id))

#read final covariates file exported to metabolon
df_cov<- read_csv("~/Documents/peach_heat_metab/metabolon_export/exp_cov_for_metabolon_dec3.csv") |>
  select(-c(12:14)) |>
  left_join(df_join, by = "id")

#rename linkage variable to join with exposures
ap_exposures<- ap_exposures |> rename(exp_id = id)

#join exposures and covariates into a dataframe for analysis
exp_cov_df<- right_join(df_cov, ap_exposures, by = c("id" = "exp_id")) |>
  mutate(across(where(is.character), as.factor))

save(exp_cov_df, file = "~/Documents/air_polln_metabolomics/analytic_dat/exp_cov_df.rda")
```




