---
title: "02_mwas"
author: "Jagadeesh Puvvula"
date: "2025-01-01"
output: pdf_document
---

#load exposure and covariates data
```{r}
load("~/Documents/air_polln_metabolomics/analytic_dat/exp_cov_df.rda")
```

#read metabolomics data
```{r}
#batch normalized and imputed for any missing metabolite intensity - as received from Metabolon
placenta_metabolome<- read_csv(paste0(data, "metabolome_batch_norm_imputed.csv")) |>
  mutate(across(2:last_col(), ~ log10(. + 0.00001))) |>
  mutate(across(2:last_col(), ~ (. - mean(.)) / sd(.)))

#annotation dictionary
df_annotations<- read_csv(paste0(data, "metabolome_annota.csv")) |>
  clean_names()

#summarizing metabolite annotations
df_summary <- df_annotations |>
  group_by(super_pathway) |>
  summarise(metabolites_count = n(),
            UNIQUE_SUB_PATHWAYS_count = n_distinct(sub_pathway),
            SUB_PATHWAYS_list = paste(sub_pathway, collapse = ", "))
```

#remove un-named metabolites for the analysis
```{r}
filtered_chem_ids <- c("PARENT_SAMPLE_NAME", df_annotations |>
                         filter(type == "NAMED") |>
                         pull(chem_id))
  
df_placenta_metabolome <- placenta_metabolome |>
  dplyr::select(any_of(as.character(filtered_chem_ids))) |>
  clean_names()
```

#drop metabolites that have identical values 
```{r}
df_placenta_metabolome_filt <- df_placenta_metabolome |>
  select(-where(~any(is.na(.))))
```


#drop 7 participants from metabolome data with missing address and dates
```{r}
filtered_placenta_metabolome <- df_placenta_metabolome_filt |>
  filter(parent_sample_name %in% exp_cov_df$parent_sample_name)

#transpose metabolite set
transposed_df <- filtered_placenta_metabolome |>
  t() |>
  as_tibble(rownames = "Variable") |>
  row_to_names(row_number = 1) |>
  rename(chem_id = 1) |>
  mutate(chem_id = sub("^x", "", chem_id))
```

#save df_update and filtered placenta_metabolome files for next steps
```{r}
save(exp_cov_df, transposed_df, df_annotations,
     file = "~/Documents/air_polln_metabolomics/analytic_dat/df_metabolome_analysis.rda")
```

#MWAS
```{r}
load("~/Documents/air_polln_metabolomics/analytic_dat/df_metabolome_analysis.rda")
```

#MWAS - loop
```{r}
appliedepi::mwas_lm_loop(feature_table = transposed_df, 
                         exp_cov_data = exp_cov_df,
                         exposures = names(exp_cov_df)[21:174],
                         covar = c("race_ethnicity", "education", "bmi", "parity",
                                   "ga_at_delivery", "tobacco_use_dur_preg", 
                                   "alcoh_use_dur_preg",
                                   "gender"), 
                         output_folder = "~/Documents/air_polln_metabolomics/result/lm_mwas",
                         mwas_file_name="placenta_mwas.csv",
                         fdr_cutoff= 0.2)
```


#load mwas results
```{r}
#beta_dir variable is based on 20% FDR cutoff
res_df<- read_csv("~/Documents/air_polln_metabolomics/result/lm_mwas/placenta_mwas.csv") |>
  clean_names() |>
  left_join(
    df_annotations |>
      select(1, 5, 6, 11, 19), 
    by = "chem_id"
  )

write_csv(res_df, "~/Documents/air_polln_metabolomics/result/lm_mwas/placenta_mwas_anno.csv")
```

#summary of results
```{r}
res_df<- read_csv("~/Documents/air_polln_metabolomics/result/lm_mwas/placenta_mwas_anno.csv")

summary_df <- res_df |>
  filter(beta_dir %in% c("negative-significant", "positive-significant")) |>
  group_by(variable, super_pathway, sub_pathway, beta_dir) |>
  summarise(count = n(), .groups = "drop") |>
  filter(count>1) |>
  mutate(exposure = case_when(
    str_starts(variable, "bc") ~ "Black carbon",
    str_starts(variable, "pm") ~ "PM2.5",
    TRUE ~ NA_character_  # Default case if neither condition is met
  ),
  exposure_window = gsub("^(pm_|bc_)", "", variable),
  exposure_window = factor(exposure_window, 
                             levels = sort(unique(as.numeric(gsub("^(pm_|bc_)", "", variable)))))
  ) |>
  mutate(super_pathway = recode(super_pathway, 
                                "Partially Characterized Molecules" = "Other",
                                "Energy" = "Other"),
         super_pathway = fct_relevel(super_pathway, 
                                     "Lipid", 
                                     "Amino Acid", 
                                     "Carbohydrate", 
                                     "Cofactors and Vitamins", 
                                     "Peptide", 
                                     "Nucleotide", 
                                     "Xenobiotics",
                                     "Other"))
```

#visualizing results - Summarized heat map
```{r}
ggplot(summary_df, aes(x = exposure_window, y = sub_pathway, fill = count)) +
  facet_grid(super_pathway~exposure, scales="free", space="free",
             switch = "both",
             labeller = label_wrap_gen(width = 15)) +
  geom_tile(color = "white") + 
  scale_fill_viridis(name = "Number of metabolites") +
  labs(x = "Gestational period (days) - Third trimester", y = "Metabolic pathways") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size=11),
        legend.position = "bottom",
        legend.text = element_text(angle = 45, hjust = 1),
        legend.title = element_text(vjust = 1),
        legend.key.height = unit(0.3, "cm"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(),
        strip.text.y = element_text(size=9, angle = 45, hjust = 0),
        strip.text.x = element_text(size=11),
        strip.placement = "outside", 
        strip.background = element_blank() )
```

```{r}
ggsave("~/Documents/air_polln_metabolomics/result/lm_mwas/mwas_pathway.tiff",
       dpi=300,
       bg="white",
       width = 9.5,
       height = 14)
```

# consolidated shapley style plot
```{r}
# Create your plot function (same as before)
create_pathway_plot <- function(data, pathway) {
  subset_data <- data %>% filter(super_pathway == pathway)
  
  ggplot(subset_data, aes(x = estimate, y = sub_pathway)) +
    geom_point(
      aes(color = case_when(
        fdr >= 0.2 ~ "lightgray",
        fdr < 0.2 & estimate > 0 ~ "blue",
        fdr < 0.2 & estimate < 0 ~ "red",
        TRUE ~ "lightgray"
      )), 
      size = 0.5, 
      position = position_jitter(height = 0.2, width = 0.1)
    ) +
    scale_color_identity() +
    labs(x = NULL, y = NULL, title = pathway) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 10),
      axis.text.y = element_text(size = 8),
      panel.spacing = unit(1, "lines")
    )
}

# Open TIFF device
tiff(
  filename = "~/Documents/air_polln_metabolomics/result/lm_mwas/mwas_lm_summmary.tiff",
  width = 12,      # Width in inches
  height = 12,      # Height in inches
  units = "in",    # Units in inches
  res = 300,       # Resolution in dpi (300 for publication quality)
  compression = "lzw"  # Compression method
)

# Create your plots
lipid_plot <- create_pathway_plot(res_df, "Lipid")

col2_plots <- list(
  create_pathway_plot(res_df, "Amino Acid"),
  create_pathway_plot(res_df, "Carbohydrate"),
  create_pathway_plot(res_df, "Nucleotide")
)

col3_plots <- list(
  create_pathway_plot(res_df, "Cofactors and Vitamins"),
  create_pathway_plot(res_df, "Energy"),
  create_pathway_plot(res_df, "Partially Characterized Molecules"),
  create_pathway_plot(res_df, "Peptide"),
  create_pathway_plot(res_df, "Xenobiotics")
)

# Create column arrangements
col2 <- arrangeGrob(grobs = col2_plots, ncol = 1)
col3 <- arrangeGrob(grobs = col3_plots, ncol = 1)

# Create final arrangement
g <- arrangeGrob(
  lipid_plot, col2, col3,
  ncol = 3,
  widths = c(1, 1, 1),
  top = textGrob("Pathway Analysis", gp = gpar(fontsize = 14)),
  bottom = textGrob("Estimate", gp = gpar(fontsize = 12)),
  left = textGrob("Sub Pathway", rot = 90, gp = gpar(fontsize = 12))
)

# Draw the plot
grid.draw(g)

# Close the device (this saves the file)
dev.off()
```

#Metabolie level plot
#load mwas results
```{r}
res_df<- read_csv("~/Documents/air_polln_metabolomics/result/lm_mwas/placenta_mwas_anno.csv")

#summary of results
summary_df <- res_df |>
  filter(beta_dir %in% c("negative-significant", "positive-significant")) |>
  mutate(prefix = ifelse(grepl("^pm_", variable), "pm_", 
                         ifelse(grepl("^bc_", variable), "bc_", NA))) |>
  filter(!is.na(prefix)) |>
  group_by(chemical_name, super_pathway, sub_pathway, prefix) |>
  summarise(count = n(), .groups = "drop") |>
  filter(count > 4)

viz_df<- res_df |>
  filter(beta_dir %in% c("negative-significant", "positive-significant"))|>
  mutate(exposure = case_when(
    str_starts(variable, "bc") ~ "Black carbon",
    str_starts(variable, "pm") ~ "PM2.5",
    TRUE ~ NA_character_  # Default case if neither condition is met
  ),
  exposure_window = gsub("^(pm_|bc_)", "", variable),
  exposure_window = factor(exposure_window, 
                             levels = sort(unique(as.numeric(gsub("^(pm_|bc_)", "", variable)))))
  ) |>
  filter(chemical_name %in% summary_df$chemical_name) |>
  mutate(
    predictor_numeric = as.numeric(gsub(".*_([0-9]+)$", "\\1", variable))
  ) |>
  mutate(super_pathway = fct_relevel(super_pathway, 
                                     "Lipid", 
                                     "Amino Acid", 
                                     "Carbohydrate", 
                                     "Cofactors and Vitamins", 
                                     "Peptide", 
                                     "Nucleotide", 
                                     "Xenobiotics"))
```

```{r}
ggplot(viz_df, aes(x = exposure_window, y = chemical_name, fill = estimate)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(
  low = scales::muted("darkred"),
  mid = "gray90", 
  high = scales::muted("darkblue"),
  midpoint = 0,
  name = "Estimate"
) +
  labs(x = "Fine particulate matter exposure during third trimester", y = "Metabolite") +
  facet_grid(super_pathway~exposure, scales="free", space="free",
             switch = "both",
             labeller = label_wrap_gen(width = 15)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        legend.position = "bottom",
        legend.text = element_text(angle = 45, hjust = 1),
        legend.title = element_text(vjust = 1),
        legend.key.height = unit(0.3, "cm"),
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank() )
```

```{r}
ggsave("~/Documents/air_polln_metabolomics/result/lm_mwas/mwas_metabolite.tiff",
       dpi=300,
       bg="white",
       width = 8,
       height = 12)
```
